Contribution Title\thanks{Supported by organization x.}

### 2020-01-27 sections/evaluation.tex
rerun experiment for data validation
sync up once again

The benchmark contains 400 instances divided into two groups of 200, new and old, representing historical instances and fresh instances in the 2019 race, respectively. I partition the old group instances into six partitions of size 30 and one partition of size 20. Each partition is then assigned to a XeonE5-2 CPU node with 16 cores (2 sockets 8 cores  and  1 thread) and 96649 MB memory. The new group is partitioned based on their contributor (e.g. Heule contributed 22 matrix multiplication instances), and each partition is assigned to a aforementioned CPU node. To speed up the experiment, we allow a CPU node to solver at most seven instances concurrently.

@@

$\IUIPMIN$ outperformed $\IUIPPURE$ in clause size.  This results agrees with our observation in Fig.~\ref{fig:t2}: $\IUIPMIN$ attempted $\IUIP$ learning more frequently, and it is more likely to succeed. Remark that the success of $\IUIP$ learning is determined by the size of the learned i-UIP clause $\iUIPClause$, and the $\IUIP$ learning frequency is also indirectly controlled by $\IUIP$'s success rate from the previous restart interval. The results indicates  $\IUIPMIN$ shortened $\iUIPClause$'s size through further resolutions of $\iUIPClause$ at non-unique implication decision level.

\begin{figure} 
\begin{center} 
\begin{tabular}{ | m{3.5cm} | m{5cm}| m{5cm} | } 
\hline
Solver & $\IUIP$ attempt rate & $\IUIP$ success rate  \\ 
\hline
$\MapleIUIPPURE$ & 16.1\% & 43.4\% \\ 
\hline
$\MapleIUIMIN$ & 28.8\% & 59.3\% \\ 
\hline
\end{tabular}
\end{center}
\caption{Compare $\IUIPPURE$ and $\IUIPMIN$ i-uip attempt rate and success rate. $\IUIPMIN$ scheme attempted $\IUIP$ more frequently, and it is more likely to successfully produce smaller $\iUIPClause$ clause .}
\label{fig:t2}
\end{figure}

@@

\begin{figure} 
\begin{center} 
\begin{tabular}{ | m{3.5cm} | m{5cm}| m{5cm} | } 
\hline
Solver & optimized proof size (MB) & relative reduction size  \\ 
\hline
$\MapleBase$ & 613.9 & 0  \\ 
\hline
$\MapleIUIPPURE$ & 487.2 & 6.90\% \\ 
\hline
$\MapleIUIMIN$ & 413.2 & 17.18\% \\ 
\hline
\end{tabular}
\end{center}
\caption{Optimized UNSAT proof comparison for 1-UIP $\IUIPPURE$ and $\IUIPMIN$. Optimized proof size measures the average absolute proof size in MB, and relative reduction size measures the average relative reduction for all UNSAT instances.}
\label{fig:t3}
\end{figure}

@@

\begin{figure} 
\begin{center}
\begin{tabular}{ | m{3.7cm} | m{4cm}| m{2cm} | m{2.75cm} |  } 
\hline
Solver & \# solved (SAT, UNSAT) & PAR-2 & Avg clause Size \\ 
\hline
SAT 2017 Winner & & & \\
$\MapleSeven$ & 232 (135, 97) & 4755.96 & 61.9  \\ 
\hline
$\MapleSeven$-i-pure & \textbf{244 (146, 98)} +12 & \textbf{4504.18} & 43.76 \\
\hline
$\MapleSeven$-i-min & 240 (144, 96) +8 & 4601.25 & \textbf{36.97} \\ 
\hline
$\MapleSeven$-i-greedy & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
\hline
$\MapleSeven$-i-inclusive & 234 (137, 97) +2 & 4718.03 & 37.96 \\
\hline
\hline
SAT 2018 Winner & & & \\
$\MapleEightShort$ & 236 (138, 98) & 4671.81 & 61.69 \\
\hline
$\MapleEightShort$-i-pure & \textbf{241} (\textbf{142, 99}) +5 & \textbf{4598.18} & 44.19 \\
\hline
$\MapleEightShort$-i-min & 236 (141, 95) +0 & 4683.92 & 38.05 \\ 
\hline
$\MapleEightShort$-i-greedy & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
\hline
$\MapleEightShort$-i-inclusive & 240 (\textbf{142}, 98) +4 & 4602.13 & \textbf{37.52} \\
\hline
\hline
SAT 2019 Winner & & & \\
$\MapleNineShort$ & 238 (140, \textbf{98}) & 4531.24 & 60.91 \\
\hline
$\MapleNineShort$-i-pure & 238 (140, \textbf{98}) +0 & 4519.08 &  43.32\\
\hline
$\MapleNineShort$-i-min & 244 (148, 96) +6 & \textbf{4409.88} & \textbf{37.43} \\
\hline
$\MapleNineShort$-i-greedy & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
\hline
$\MapleNineShort$-i-inclusive & \textbf{245} (\textbf{150}, 95) +7 & 4425.99 & 37.60 \\
\hline
\hline
SAT 2019 Competitor & & &\\
$\expSATShort$ & 237 (137, 100)  & 4628.96 & 63.19 \\
\hline
$\expSATShort$-i-pure & 235 (136, 99) -2  & 4668.96 & 48.26 \\
\hline
$\expSATShort$-i-min & 241 (143, 98) +4 & 4524.28 & 46.29 \\ 
\hline
$\expSATShort$-i-greedy & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
\hline
$\expSATShort$-i-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33} \\
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of 1-UIP, $\IUIPPURE$. $\IUIPMIN$, $\IUIPGreedy$ and $\IUIPActive$ on SAT2019 race main track.}
\label{fig:t5}
\end{figure}

\nf{Benchmark results from running $\expSAT$ are harder to produce due to the randomness caused by the solver's strategy for selecting branching heuristics: the solver runs 2500s with LBR and the remaining 2500s with VSIDS. This is problematic for reproducing results because two runs of the same solver can be in different states at 2500s mark, and switching heuristics at different states will cause further divergence. The result can be quiet impactful (+- 2 instances at most for i-uip-active). For reproducible benchmark results, should we turn off the 2500s heuristic switching? Should we add a new section for the reproducible result? or remove the existing ones?}

Table~\ref{fig:t5} shows the benchmark result of $\IUIP$ configurations on different solvers. All four configurations of $\IUIP$ outperformed 1-UIP learning for the SAT 2017 race winner, $\MapleSeven$. More specifically, $\IUIPPURE$, $\IUIPMIN$, $\IUIPGreedy$ and $\IUIPActive$ solved 12, 8, 5 and 2 more instances, respectively, whiling producing smaller clauses. $\IUIPPURE$ solved more UNSAT and SAT instances while other configurations improved on solving SAT instances.  The improvement of $\IUIP$ is more significant on $\MapleSeven$ than on $\MapleBase$ for both solved instances and clause size reduction. This may suggests that $\IUIP$ and the recent learnt clause minimization approach~\cite{} synergies well because i-UIP clauses are shorter with more common literals which allows vivification~\cite{} to prune literals more aggressively through more unit propagation.

We observed significant improvement of $\IUIP$ for the SAT 2018 winner, $\MapleEightShort$. Three out of four $\IUIP$ configurations outperformed 1-UIP by 5, 4 an 4 instances, respectively. $\MapleEightShort$ improved from $\MapleSeven$ by using chronological backtracking (CB) for long distance backtracks. Therefore, we used $\IUIP$-CB extension for all $\IUIP$ configurations. The results indicates that the improvement of $\IUIP$ is slightly shadowed by the adoption of CB. More specifically, we believe CB prevents decision levels from being compressed through the process of backtracking and literal assertion. Comparing to 1-UIP clause, the shorter i-UIP clauses can bring related literals closer with shorter implications, and consequently produces less and smaller decision levels through  backtracking. However, since CB discourages long distance backtracking, the effect of learning shorter clauses is shadowed until a full restart.  We believe we have observed an interesting interaction effect that shows the limitations of both CB and $\IUIP$ learning for future research.

$\IUIP$ learning schemes showed significant improvement for $\MapleNineShort$. Three out of four configurations improved solved instances by 6, 5 and 7 instances, respectively. $\MapleNineShort$ uses duplicate learning (DL) to prioritize clauses that are learned multiple times. $\IUIP$ and DL didn't synergies well possibly because i-uip clauses are less likely to be duplicated. We observed that $\IUIPMIN$ ,in average, added 12\% less duplicated clauses into the core clause database. One possible explanation is that the conflict blocked by a single 1-uip clauses can be alternatively blocked by one of many i-uip clauses. Therefore, the same 1-uip clause can be reduced to different i-uip clauses depending on the solver's assertion trail.

We observed significant improvement of $\IUIP$ for $\expSATShort$, three out of four configurations of $\IUIP$ significantly outperformed 1-UIP learning by 4, 7 and 8 more instances, respectively, while producing smaller clauses. We expect all three configurations of $\IUIP$ to solve the similar amount of instances with close PAR-2 scores because the additional random walk exploration allows the learning schemes to partially sidestep the activity problem; hence, the learning adjustment for variable activities should have less impact on the solver's performance. However, we observed that both $\IUIPGreedy$ and $\IUIPActive$ outperformed the default i-UIP learning scheme. One possible explanation is that $\IUIPGreedy$ and $\IUIPActive$ schemes could easily overcompensate variable activities for i-UIP clauses, and $\expSATShort$'s random walk exploration could use future search information to mitigate the negative effects of our overcompensation.

>>>>

The benchmark contains 400 instances divided into two groups of 200, new and old, representing historical instances and fresh instances in the 2019 race, respectively. We randomly partition the old group instances into six partitions of size 30 and one partition of size 20. Each partition is then assigned to a XeonE5-2 CPU node with 16 cores (2 sockets 8 cores  and  1 thread) and 96649 MB memory. The new group is partitioned based on their contributor (e.g. Heule contributed 22 matrix multiplication instances), and each partition is assigned to a aforementioned CPU node. To speed up the experiment, we allow a CPU node to solver at most seven instances concurrently.

@@

$\IUIPMIN$ outperformed $\IUIPPURE$ in clause size.  This results agrees with our observation in Fig.~\ref{fig:t2}: $\IUIPMIN$ attempted $\IUIP$ learning more frequently, and it is more likely to succeed.

\begin{figure} 
\begin{center} 
\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
\hline
Solver & $\IUIP$ attempt rate & $\IUIP$ success rate  \\ 
\hline
$\MapleIUIPPURE$ & 16.1\% & 43.4\% \\ 
\hline
$\MapleIUIMIN$ & 28.8\% & 59.3\% \\ 
\hline
\end{tabular}
\end{center}
\caption{Compare $\IUIPPURE$ and $\IUIPMIN$ i-uip attempt rate and success rate. $\IUIPMIN$ scheme attempted $\IUIP$ more frequently, and it is more likely to successfully produce smaller $\iUIPClause$ clause .}
\label{fig:t2}
\end{figure}

@@

\begin{figure} 
\begin{center} 
\begin{tabular}{ | m{3.5cm} | m{5cm}| m{3.5cm} | } 
\hline
Solver & optimized proof size (MB) & relative reduction size  \\ 
\hline
$\MapleBase$ & 613.9 & 0  \\ 
\hline
$\MapleIUIPPURE$ & 487.2 & 6.90\% \\ 
\hline
$\MapleIUIMIN$ & 413.2 & 17.18\% \\ 
\hline
\end{tabular}
\end{center}
\caption{Optimized UNSAT proof comparison for 1-UIP $\IUIPPURE$ and $\IUIPMIN$. Optimized proof size measures the average absolute proof size in MB, and relative reduction size measures the average relative reduction for all UNSAT instances.}
\label{fig:t3}
\end{figure}

@@

\begin{figure} 
\begin{center}
\begin{tabular}{ | m{3.7cm} | m{4cm}| m{2cm} | m{2.75cm} |  } 
\hline
Solver & \# solved (SAT, UNSAT) & PAR-2 & Avg clause Size \\ 
\hline
SAT 2017 Winner & & & \\
$\MapleSeven$ & 232 (135, 97) & 4755.96 & 61.9  \\ 
\hline
$\MapleSeven$-i-pure & \textbf{244 (146, 98)} +12 & \textbf{4504.18} & 43.76 \\
\hline
$\MapleSeven$-i-min & 240 (144, 96) +8 & 4601.25 & \textbf{36.97} \\ 
\hline
$\MapleSeven$-i-greedy & 237 (140, 97) +5 & 4678.434 & 43.62 \\ 
\hline
$\MapleSeven$-i-inclusive & 234 (137, 97) +2 & 4718.03 & 37.96 \\
\hline
\hline
SAT 2018 Winner & & & \\
$\MapleEightShort$ & 236 (138, 98) & 4671.81 & 61.69 \\
\hline
$\MapleEightShort$-i-pure & \textbf{241} (\textbf{142, 99}) +5 & \textbf{4598.18} & 44.19 \\
\hline
$\MapleEightShort$-i-min & 236 (141, 95) +0 & 4683.92 & 38.05 \\ 
\hline
$\MapleEightShort$-i-greedy & 240 (141, \textbf{99}) +4 & 4626.99 & 41.16 \\
\hline
$\MapleEightShort$-i-inclusive & 240 (\textbf{142}, 98) +4 & 4602.13 & \textbf{37.52} \\
\hline
\hline
SAT 2019 Winner & & & \\
$\MapleNineShort$ & 238 (140, \textbf{98}) & 4531.24 & 60.91 \\
\hline
$\MapleNineShort$-i-pure & 238 (140, \textbf{98}) +0 & 4519.08 &  43.32\\
\hline
$\MapleNineShort$-i-min & 244 (\textbf{148}, 96) +6 & \textbf{4419.84} & \textbf{36.88} \\
\hline
$\MapleNineShort$-i-greedy & 243 (146, 97) + 5 & 4476.73 & 40.65 \\
\hline
$\MapleNineShort$-i-inclusive & \textbf{243} (\textbf{148}, 95) +5 & 4455.76 & 37.02 \\
\hline
\hline
SAT 2019 Competitor & & &\\
$\expSATShort$ & 237 (137, 100)  & 4628.96 & 63.19 \\
\hline
$\expSATShort$-i-pure & 235 (136, 99) -2  & 4668.96 & 48.26 \\
\hline
$\expSATShort$-i-min & 241 (143, 98) +4 & 4524.28 & 46.29 \\ 
\hline
$\expSATShort$-i-greedy & 244 (143, \textbf{101}) +7 & \textbf{4460.92} & 47.25 \\
\hline
$\expSATShort$-i-inclusive & \textbf{245} (\textbf{146}, 99) +8 & 4475.76 & \textbf{45.33} \\
\hline
\end{tabular}
\end{center}
\caption{Benchmark results of 1-UIP, $\IUIPPURE$. $\IUIPMIN$, $\IUIPGreedy$ and $\IUIPActive$ on SAT2019 race main track.}
\label{fig:t5}
\end{figure}

Table~\ref{fig:t5} shows the benchmark result of $\IUIP$ configurations on different solvers. All four configurations of $\IUIP$ outperformed 1-UIP learning for the SAT 2017 race winner, $\MapleSeven$. More specifically, $\IUIPPURE$, $\IUIPMIN$, $\IUIPGreedy$ and $\IUIPActive$ solved 12, 8, 5 and 2 more instances, respectively, whiling producing smaller clauses. $\IUIPPURE$ solved more UNSAT and SAT instances while other configurations improved on solving SAT instances.  The improvement of $\IUIP$ is more significant on $\MapleSeven$ than on $\MapleBase$ for both solved instances and clause size reduction. This may suggests that $\IUIP$ and the recent learnt clause minimization approach~\cite{} synergies well because i-UIP clauses are shorter with more common literals which allows vivification~\cite{} to prune literals more aggressively through unit propagation.

We observed significant improvement of $\IUIP$ for the SAT 2018 winner, $\MapleEightShort$. Three out of four $\IUIP$ configurations outperformed 1-UIP by 5, 4 an 4 instances, respectively. $\MapleEightShort$ improved from $\MapleSeven$ by using chronological backtracking (CB) for long distance backtracks. Therefore, we used $\IUIP$-CB extension for all $\IUIP$ configurations. The results indicates that the improvement of $\IUIP$ is slightly shadowed by the adoption of CB. More specifically, we believe CB prevents decision levels from being compressed through the process of backtracking and literal assertion. The shorter i-UIP clauses can bring related literals closer, which in turns compresses the assertion stack and the decision levels. However, since CB discourages long distance backtracking, the effect of learning shorter clauses is shadowed until a full restart.  We believe we have observed an interesting interaction effect that shows the limitations of both CB and $\IUIP$ learning for future research.

$\IUIP$ learning schemes showed significant improvement for $\MapleNineShort$. Three out of four configurations improved solved instances by 6, 5 and 5 instances, respectively. $\MapleNineShort$ uses duplicate learning (DL) to prioritize clauses that are learned multiple times. $\IUIP$ and DL didn't synergies well possibly because i-UIP clauses are less likely to be duplicated. We observed that $\IUIPMIN$ ,in average, added 12\% less duplicated clauses into the core clause database than 1-UIP. One possible explanation is that $\IUIP$ learning can reduce an 1-UIP clause to different i-UIP clauses for different assertion trails. Therefore, the solver is less likely to learn the same i-UIP clauses.

We observed significant improvement of $\IUIP$ for $\expSATShort$, three out of four configurations of $\IUIP$ significantly outperformed 1-UIP learning by 4, 7 and 8 more instances, respectively, while producing smaller clauses. We expect all three configurations of $\IUIP$ to solve the similar amount of instances with close PAR-2 scores because the additional random walk exploration allows the learning schemes to partially sidestep the activity problem; hence, the learning adjustment for variable activities should have less impact on the solver's performance. However, we instead, observed that both $\IUIPGreedy$ and $\IUIPActive$ outperformed the default i-UIP learning scheme. One possible explanation is that $\IUIPGreedy$ and $\IUIPActive$ schemes could easily overcompensate variable activities for i-UIP clauses, and $\expSATShort$'s random walk exploration could use future search information to mitigate the negative effects of our overcompensation.
